# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
unzip(zipfile="./activity.zip")
data <- read.csv(file="./activity.csv")
data$date <- as.Date(data$date)
summary(data)
```


## What is mean total number of steps taken per day?
```{r}
library(ggplot2)
qplot(steps, data=data, geom='histogram', binwidth = 40)
#not clear if we have to cut by day...
#mn <- aggregate(steps ~ cut(data$date, "day"), data=data, FUN=mean, na.rm=T)
#md <- aggregate(steps ~ cut(data$date, "day"), data=data, FUN=median, na.rm=T)
```

Mean steps per day is `r mean(data$steps, na.rm=T)`.
Median steps per day is `r median(data$steps, na.rm=T)`.


## What is the average daily activity pattern?
```{r}
y <- list(as.factor(data$interval))
mean.by.5 <- aggregate(data$steps, y, mean, na.rm=T)
names(mean.by.5) <- c("interval", "mn")

plot(mean.by.5$interval, mean.by.5$mn, type = "n")
lines(mean.by.5$interval, mean.by.5$mn)
max.point <- mean.by.5[which(mean.by.5$mn == max(mean.by.5$mn)), ]
points(max.point, pch=4, cex=2, col="red")
```

5'-interval with max steps in avg across all days: `r max.point$interval` with `r max.point$mn` steps in average.

## Imputing missing values
Number of rows with NAs is `r sum(!is.na(data$steps))`.

We're going to fill the gaps following the simplest way:
missing value for a the `x`th 5 minutes interval in a day will got the mean of all the non `r NA`s values for this interval.

```{r}
data$steps.completed <- data$steps
for (m in levels(mean.by.5$interval)) {
  idx  <- which(is.na(data$steps.completed) & data$interval==m)
  init <- data[idx, ]$steps.completed
  v <- mean.by.5[mean.by.5$interval == m, ]$mn
  data[idx, ]$steps.completed <- v
  after <- data[idx, ]$steps.completed
}
```

Now the number of NAs is `r sum(is.na(data$steps.completed))`.


```{r}
qplot(steps.completed, data=data, geom='histogram')
#not clear if we have to cut by day...
#mn <- aggregate(steps ~ cut(data$date, "day"), data=data, FUN=mean, na.rm=T)
#md <- aggregate(steps ~ cut(data$date, "day"), data=data, FUN=median, na.rm=T)
```

Mean steps per day is `r mean(data$steps.completed, na.rm=T)`.
Median steps per day is `r median(data$steps.completed, na.rm=T)`.

The mean and median doesn't change most probably because the data is skewed at 0 steps (median invariant) and we used the mean to fulfill the missing values (the mean invariant).


## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
data$weekend <- weekdays(data$date) %in% c("Saturday", "Sunday")
data[data$weekend == T, ]$weekend <- "Weekend"
data[data$weekend == F, ]$weekend <- "Weekday"
gs <- list(as.factor(data$interval), as.factor(data$weekend))
completed.mean.by.5 <- aggregate(data$steps, gs, mean, na.rm=T)
names(completed.mean.by.5) <- c("interval", "weekend", "mn")
xyplot(mn ~ as.numeric(as.vector(completed.mean.by.5$interval)) | weekend, type="l", data=completed.mean.by.5, layout=c(1, 2), xlab="interval in day", ylab="mean # of steps")
```